# Simulations {#sec-sim}


:::{.callout-warning}

### Warning

The estimations shown in this chapter were run outside the compilation of this document. The codes are shown here, but they are not evaluated to create this HTML page.

In total, we consider 6 different configuration for the data shown in [Chapter -@sec-sim-functions]. The first 3 configurations take about 2h30 to run each, while the 3 next take about 12h each.

:::

:::{.callout-note collapse=true}

## Required packages

Some required packages:
```{r load-packages}
library(tidyverse)
library(pbapply)
library(hdm)
library(randomForest)
library(xgboost)
library(sandwich)
library(philentropy)
library(parallel)
library(gmish)
library(ebal)
library(MASS)

# To install gmish:
# devtools::install_git("gweissman/gmish")
```

:::

The functions presented in [Chapter -@sec-sim-functions] are all stored in scripts in the `scripts/simulation` directory. We load those.

```{r}
# Source modular components
source("../scripts/simulation/01_dgp.R")
source("../scripts/simulation/02_learners_g.R")
source("../scripts/simulation/03_learners_m.R")
source("../scripts/simulation/04_calibration.R")
source("../scripts/simulation/05_metrics.R")
source("../scripts/simulation/06_causal-estimators.R")
source("../scripts/simulation/07_simulation_helper.R")
```


## Parameter Grid Setup

We define a grid with varying values for some parameters that influence either the data generating process ($n$, $p$, $R_d^2$) or whether propensity scores are clipped. For the latter, note that we set a very small value, which corresponds to no clipping. A value of 0 may produce issues when computing the AIPW estimator or the ATE. Also, note that for Beta calibration, the number of drawn data needs to be quite large. We tried $n=100$ and $n=200$ but this leads to issued when trying to calibrate with Beta calibration.

```{r define-grid}
grid <- expand_grid(
  n    = c(2000),
  p    = c(20, 200),
  Rd2  = c(0.2, 0.5, 0.8),
  clip = 1e-12 # 1e-12 corresponds to virtually no clipping
)
grid
```


The results are stored in the `output/simul-klassen/` folder.
```{r define-output_dir}
output_dir <- "../output/simul-klassen/"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
```

## Parallel Setup

For each configuration of the generated data, we run 1,000 replications of the Monte Carlo simulations.
```{r define-nb_simul}
nb_simul <- 1000
```


Since these replications are independent from one another, they can be run in parallel. The helper functions need to be sent to the clusters.



```{r parallel-setup, eval=FALSE}
ncl <- max(1, detectCores() - 1)
cl  <- makeCluster(ncl)
# Guarantee cluster shutdown even if an error occurs mid-script
on.exit(stopCluster(cl), add = TRUE)

# Load dependencies across all worker nodes
clusterEvalQ(cl, {
  library(tidyverse)
  library(hdm)
  library(randomForest)
  library(xgboost)
  library(sandwich)
  library(philentropy)
}) |> invisible()

# Export custom helper and estimation functions
functions_to_export <- c(
  "simul",
  # Data Generation
  "gen_data_belloni", "format_data_belloni",
  # Learners g
  "estim_g", "estim_g_ols", "predict_g_ols",
  "estim_g_lasso", "predict_g_lasso",
  "estim_g_rf", "predict_g_rf",
  "estim_g_lgbm", "predict_g_lgbm",
  # Learners m
  "estim_m",
  "estim_m_logit", "predict_m_logit",
  "estim_m_rf_classif", "predict_m_rf_classif",
  "estim_m_rf_classif", "predict_m_rf_classif",
  "estim_m_rf_reg", "predict_m_rf_reg",
  "estim_m_lgbm", "predict_m_lgbm",
  # Calibration
  "calibrate", "nuisance_estim", "alg_1_uncalibrated",
  "alg_3_cross_fitted_calib", "calibration_curve",
  # Metrics
  "compute_ece", "kullback_leibler",
  # Causal Estimators
  "compute_ate_reg", "compute_ate_ipw", "compute_ate_aipw",
  "compute_ate_ipw_normalized", "compute_ate_aipw_normalized",
  "compute_ate_eqb_reg", "compute_ate_eqb_ipw", "compute_ate_eqb_aipw",
  "compute_ate_folds",
  "qtmax", "group_average_treatment_effect",
  # Simulation Helpers
  "compute_simul_metrics", "compute_simul_gate", "resolve_clipping_and_calib"
)

clusterExport(cl, functions_to_export)
```


## Main Simulation Loop

```{r, eval=FALSE}
cat(sprintf("Starting Monte Carlo simulations (%d combinations across %d cores)...\n\n",
            nrow(grid), ncl))

for (i_grid in 2:nrow(grid)) {
  cat(
    sprintf(
      "## Simulation %d/%d (n=%d, p=%d, Rd2=%.1f, clip=%g)\n",
      i_grid, nrow(grid), grid$n[i_grid], grid$p[i_grid],
      grid$Rd2[i_grid], grid$clip[i_grid])
  )

  dgp_params <- list(
    n   = grid$n[i_grid],
    p   = grid$p[i_grid],
    Rd2 = grid$Rd2[i_grid],
    Ry2 = 0.5,
    rho = 0.5
  )

  clipping_threshold <- grid$clip[i_grid]

  # Export current parameter state to worker nodes
  clusterExport(cl, c("dgp_params", "clipping_threshold"), envir = environment())

  # Run replicates in parallel
  res_tmp <- pblapply(1:nb_simul, FUN = function(seed) {
    simul(
      dgp_name           = "belloni",
      dgp_params         = dgp_params,
      clipping_threshold = clipping_threshold,
      n_folds            = 5,
      seed               = seed
    )
  }, cl = cl)

  # Save individual grid output iteration
  save(res_tmp, file = str_c(output_dir, i_grid, ".rda"))
}

cat("\nSimulations completed successfully.\n")
```




